---
title: "Praccomp Final Project"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---
# Setting Working Directory
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Grace/praccomp2024/final_project")
```

# Loading necessary packages
```{r}
#Rank abundance curves for trematode diversity data
library(vegan)
library(BiodiversityR)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(tidyverse)
#Oyster density plots
library(ggplot2)
library(MASS)
```

# Trematode Diversity Analyses

## Loading in the data
```{r}
#Loading in the data
BIOR_5YR <- read.csv("BIOR_5YR.csv")
BIOR_5YR_env <- read.csv("BIOR_5YR_env.csv")

BIOR_Pre <- read.csv("BIOR_Pre.csv")
BIOR_1YR <- read.csv("BIOR_1YR.csv")
BIOR_PRE_1YR_env <- read.csv("BIOR_PRE_1YR_env.csv")
```

## Changing Site Type to a Factor in the Environmental data
```{r}
BIOR_5YR_env$site_type <-as.factor(BIOR_5YR_env$site_type)
class(BIOR_5YR_env$site_type)

BIOR_PRE_1YR_env$site_type <-as.factor(BIOR_PRE_1YR_env$site_type)
class(BIOR_PRE_1YR_env$site_type)
```


## Plots within the BiodiversityR package functions
```{r}
#5YR - quick plot 

rankabuncomp(BIOR_5YR, y=BIOR_5YR_env, factor='site_type', 
    scale='logabun', legend=FALSE)

#1YR - quick plot 

rankabuncomp(BIOR_1YR, y=BIOR_PRE_1YR_env, factor='site_type', 
    scale='logabun', legend=FALSE)

#Pre - quick plot

rankabuncomp(BIOR_Pre, y=BIOR_PRE_1YR_env, factor='site_type', 
    scale='logabun', legend=FALSE)
```


## Defining a theme to use in the ggplot curves
```{r}
#This is an online theme that is provided along with rank-abundance curve code, I included it here to help enhance the visual produced by the rank abundance curve plots.
BioR.theme <- theme(
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.line = element_line("gray25"),
        text = element_text(size = 12),
        axis.text = element_text(size = 10, colour = "gray25"),
        axis.title = element_text(size = 14, colour = "gray25"),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.key = element_blank())
```


## Making Rank-Abundance Curves using ggplot
```{r}
#fancier plots

#PRE

fancy_Pre <- rankabuncomp(BIOR_Pre, y=BIOR_PRE_1YR_env, factor='site_type', scale='logabun',  return.data=TRUE, legend=FALSE)

fancy_Pre

fancy_Pre_plot <- ggplot(data=fancy_Pre, aes(x = rank, y = log(abundance))) +
    scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(aes(colour=Grouping), size=1) + 
    geom_point(aes(colour=Grouping, shape=Grouping), size=5, alpha=0.7) +
    geom_text_repel(data=subset(fancy_Pre, labelit == TRUE), 
        aes(colour=Grouping, label=species), 
        angle=45, nudge_x=1, nudge_y=1, show.legend=FALSE) + 
  BioR.theme +
    labs(x = "Rank", y = "Abundance", title= "Pre-restoration (2017-2018)", colour = "Site Type", shape = "Site Type") +
  scale_color_manual(values=c("#666666","#669933"))


fancy_Pre_plot

#1 YR

fancy_1YR <- rankabuncomp(BIOR_1YR, y=BIOR_PRE_1YR_env, factor='site_type', scale='logabun',return.data=TRUE, legend=FALSE)

fancy_1YR_plot <- ggplot(data=fancy_1YR, aes(x = rank, y = log(abundance))) +
    scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(aes(colour=Grouping), size=1) + 
    geom_point(aes(colour=Grouping, shape=Grouping), size=5, alpha=0.7) +
    geom_text_repel(data=subset(fancy_1YR, labelit == TRUE), 
        aes(colour=Grouping, label=species), 
        angle=45, nudge_x=1, nudge_y=1, show.legend=FALSE) + 
  BioR.theme +
    labs(x = "Rank", y = "Abundance", title= "One year Post-restoration (2018-2019)", colour = "Site Type", shape = "Site Type") +
  scale_color_manual(values=c("#666666","#669933"))



fancy_1YR_plot


#5 YR 

fancy_5YR_nolog <- rankabuncomp(BIOR_5YR, y=BIOR_5YR_env, factor='site_type', scale='logabun', return.data=TRUE, legend=FALSE)

fancy_5YR_plot2 <- ggplot(data=fancy_5YR_nolog, aes(x = rank, y = log(abundance))) + 
    scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(aes(colour=Grouping), size=1) +
    geom_point(aes(colour=Grouping, shape=Grouping), size=5, alpha=0.7) +
    geom_text_repel(data=subset(fancy_5YR_nolog, labelit == TRUE), 
        aes(colour=Grouping, label=species), 
        angle=45, nudge_x=1, nudge_y=1, show.legend=FALSE) +
    BioR.theme +
    labs(x = "Rank", y = "Abundance", title = "Five years Post-restoration (2023-2024)", colour = "Site Type", shape = "Site Type") + scale_color_manual(values=c("#666666","#669933"))



fancy_5YR_plot2 
```

## Combining the RA curves to one figure
```{r, fig.width=20, fig.height=6}
fancy_Pre_plot + fancy_1YR_plot + fancy_5YR_plot2
```

# Oyster density data analyses

## Loading the data
```{r}
Newoysdens1 <- read.csv("Newoysdens1.csv")
```

#Plotting the oyster density data
```{r}
stackedoys1<-ggplot(Newoysdens1, aes(x=as.factor(Method), y=Oysdens, fill=Shoreline)) + 
  geom_boxplot()+ 
  xlab("Restoration Method")+ ylab("Oyster Densities (Oysters/m2)") + labs(fill="Shoreline and Age")
  
stackedoys1 + scale_fill_manual(values=c("#BDD7EE","#3B4F23","#317ABD","#82AD4D")) + theme(panel.background = element_rect(fill = 'lightyellow3'))+
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=15,face="bold"),
        legend.text = element_text(size=12),
        legend.title = element_text(size=13))
```

## Checking the distribution of the data and log-transforming for GLMs
```{r}
hist(Newoysdens1$Counts)
log_counts<-log(Newoysdens1$Counts)

noyster<-cbind(Newoysdens1,log_counts)
hist(noyster$log_counts)
```

#Preliminary Analysis with Generalized Linear Models (GLMs)
```{r}
# I ran multiple models and this one appeared to have the best AIC score.
GLM_combo<-glm(Counts~Shoreline+Method, data = noyster, family = poisson)
summary(GLM_combo)
```

#Future Analyses
```{r}
#Linear mixed effects model with time, season, and water quality as fixed effects
#In addition to oyster densities, I want to look at how these effects may impact the reef footprint, marsh edge location, and marsh stem density/heights.
```



